---
title: "FZJ WWTP ATS Pilot"
author: "Dean Calahan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r FWAP_setup, echo = FALSE, include = TRUE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE)
knitr::opts_knit$set(root.dir = "../../")
```

# FZJ WWTP ATS Pilot

Knitting this file additionally builds a cleaned data set if the directory ./Data/ has been deleted. The source data are read from xlsx files contained in the directory *Data/xlsx/[year]/*. Raw data files are located in *Research/Data/* and were copied there from their orginal sources on the FZJ campus network. Raw data comprises .xlsx, .xls, and .HOBO files. Non-.xlsx files were converted to .xlsx within their original source directories on the FZJ campus network, using Microsoft Excel v or HOBOWare v on Windows v.
                
## Project Settings
```{r FWAP_proj_settings}
source("Code/R/Settings.R")
```

## Data integrity
```{r FWAP_copy_files}
# If ./Data/ exists, assume it is valid, else copy the original data from Research/Data/
CheckData()
```

## Water quality and biomass data  
### Load the raw data
```{r FWAP_load_WC2018}
# Water chemistry and biomass productivity
treatment_2018 <- WaterChemistryBiomass2018(
    file.path(data_dir, "xlsx", "2018", "ATS Treatment_last version.xlsx"))
```
```{r FWAP_load_WC2019}
treatment_2019 <- WaterChemistryBiomass2019(
    file.path(data_dir, "xlsx", "2019", "Treatment Data.xlsx"))
```
```{r FWAP_load_BC2018}
# Biomass Elemental composition. In 2018 it was all in one spreadsheet
biomass_2018 <- BiomassCompositionData2018(
    file.path(data_dir, "xlsx", "2018", "Analysen ZEA-3", "Analysen ZEA-3.xlsx")
    )
```
```{r FWAP_load_BC2019}
# Biomass Elemental composition. In 2019 it was in two spreadsheets
biomass_2019 <- BiomassCompositionData2019(
    file.path(data_dir, "xlsx", "2019", "191206_Ergebnisse_ICPOES.xlsx"),
    file.path(data_dir, "xlsx", "2019", "auf191206_EA.XLSX")
    )
```

### Display raw water chemistry and biomass data
```{r FWAP_tbl_WC_2018}

knitr::kable(
  treatment_2018 %>% select(datetime, before, PO4P, TN, pH, frozen, assaydate, wet_biomass, dry_biomass),
  caption = 'Water Treatment 2018 A'
)
knitr::kable(
  treatment_2018 %>% select(solids, ZEA, obsdate, person),
  caption = 'Water Treatment 2018 B'
)
```
```{r FWAP_tbl_WC_2019}
knitr::kable(
  treatment_2019 %>% select(datetime, before, PO4P, TN, pH, frozen, assaydate, wet_biomass, dry_biomass),
  caption = 'Water Treatment 2019 A'
)
knitr::kable(
  treatment_2019 %>% select(solids, ZEA, obsdate, person),
  caption = 'Water Treatment 2019 B'
)
```
```{r FWAP_tbl_BC_2018}
knitr::kable(
  biomass_2018,
  caption = 'Biomass 2018'
)
```
```{r FWAP_tbl_BC_2019}
knitr::kable(
  biomass_2019,
  caption = 'Biomass 2019'
)
```

### Assemble raw water chemistry and biomass data into clean data sets
```{r FWAP_WC_2019}
treatment <- treatment_2018 %>%
    select(datetime, before, PO4P, TN, pH, frozen, assaydate, obsdate) %>%
    bind_rows(
        treatment_2019 %>% select(datetime, before, PO4P, TN, pH, frozen, assaydate, obsdate)
    )

harvest <- treatment_2018 %>%
    filter(!is.na(dry_biomass)) %>%
    select(datetime, wet_biomass, dry_biomass) %>%
    bind_rows(
        treatment_2019 %>%
            filter(!is.na(dry_biomass)) %>%
            select(datetime, wet_biomass, dry_biomass)
    )

biomass <- biomass_2018 %>%
    select(date, element, wtprop, wtpropsd) %>%
    rename(percent = wtprop, sd = wtpropsd) %>%
    mutate(type = "harvest") %>%
    bind_rows(
        biomass_2019 %>%
            select(element, pct, sd) %>%
            mutate(type = "mix", date = ymd("2019-06-12", tz = "Europe/Berlin"))
    )
```
## HOBO data
Work in progress
```{r FWAP_HOBO_data}
```

## Irradiance data
Work in progress
```{r FWAP_irradiance_data}
```

## Create Folders and Files for Dataset
```{r FWAP_save_treatment}
treatment %>% write.csv(file.path(data_dir, "csv", "treatment.csv"))
harvest %>% write.csv(file.path(data_dir, "csv", "harvest.csv"))
biomass %>% write.csv(file.path(data_dir, "csv", "biomass.csv"))
```