---
title: 'Forschungszentrum-Jülich ATS pilot 2018'
author:
- Dean Calahan^[Smithsonian Institution, Forschungszentrum Jülich]
- Isabel Meuser and Ladislav Nedbal^[Forschungszentrum Jülich]
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
urlcolor: blue
references:
- id: schreiber2017
  title: Growth of algal biomass in laboratory and in large-scale algal photobioreactors in the temperate climate of western Germany
  author:
  - family: Schreiber
    given: Christina
  - family: Behrendt
    given: Dominik
  - family: Huber
    given: Gregor
  - family: Pfaff
    given: Christian
  - family: Widzgowski
    given: Janka
  - family: Ackermann
    given: Bärbel
  - family: Müller
    given: Andreas
  - family: Zachleder
    given: Vilém
  - family: Moudríková
    given: Šárka
  - family: Mojzeš
    given: Peter
  container-title: Bioresource Technology
  volume: 234
  URL: 'https://doi.org/10.1016/j.biortech.2017.03.028'
  DOI: 10.1016/j.biortech.2017.03.028
  publisher: Elsevier
  page: 140-149
  type: article-journal
  issued:
    year: 2017
    month: 6
---

```{r header}
# # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # #
#
# Copyright 2018-2019 by Forschungszentrum-Jülich (FZJ)
#
# This file is part of digital object 10.5281/zenodo.2593844, "FZJ ATS Pilot 2018"
# (FZJAP2018), a data set documenting the results of an algal turf scrubbing pilot
# project performed 15.07-15.09 2018, treating secondary sewage at FZJ's campus 
# wastewater treatment plant.
#
# The software component of FZJAP2018 is open access: you can redistribute it and/or
# modify it under the terms of the GNU General Public License as published by the
# Free Software Foundation, either version 3 of the License, or (at your option)
# any later version.
#
# The software component of FZJAP2018 is distributed in the hope that it will be
# useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
# or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License for more
# details.
#
# You should have received a copy of the GNU General Public License along with the
# software component of FZJAP2018. If not, see <http://www.gnu.org/licenses/>.
#
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, cache = TRUE, fig.height=3)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
# # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # #
# Execute one of the following commands to knit, rather than use the knit button.
# rmarkdown::render("Code/Rmd/FZJAP2018.Rmd", output_dir="Drafts", output_format = c("pdf_document"))
# rmarkdown::render("Code/Rmd/FZJAP2018.Rmd", output_dir="Drafts", output_format = c("html_document",  "pdf_document", "word_document"))
# # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # #
```
```{r initialize}
#
# [todo] Determine the source of errors when using the following YAML header to create
# a Word document. Perhaps setting root.dir is causative.
#
# output 
#   html_document: default
#   pdf_document: default
#   word_document: default
# knit: (function(inputFile, encoding) {
#   rmarkdown::render(inputFile, encoding = encoding,
#   output_dir = "output", output_format = "all") })
source("Code/R/Settings.R")
source("Code/R/FZJAP2018.R")
```
# Introduction  
Here we provide [**cleaned data, in the form of csv files**](https://doi.org/10.5281/zenodo.2593844),
from the 2018 ATS pilot
at Forschungszentrum-Jülich conducted by the Institute for Bio- and Geosciences's
Plant Sciences: Alternative Biomass department. Biomass composition data were obtained 
by standard analyses and recorded in the spreadsheet `Analysen ZEA-3.xlsx`. Biomass
productivity data were recorded manually in the spreadsheet `ATS Treatment.xlsx`.
Temperature and irradiance data were obtained from HOBO data loggers and exported
as csv files. Data was cleaned by knitting `FZJAP2018.Rmd`, which, using functions
defined in `FZJAP2018.R` rearranges, renames, or or discards cells or observations.
Four csv files are created, recording biomass composition and productivity, logged
temperature and irradiance, and water chemistry.

# Methods
*Floway Configuration*. A 0.54 m × 2.5 m algal turf scrubber (ATS) floway and headworks was assembled from
local home improvement store materials (Bauhaus, Belp, Switzerland; Hornbach Baumarkt
AG, Bornheim, Germany) and lined with surplus growth substratum originally used
for a thin layer algae spray reactor [@schreiber2017]. The floway was operated
at the Forschungszentrum-Jülich (FZ-J) campus wastewater treatment plant (WWTP) 
from August 15th to October 22nd, 2018. Influent was recirculated in batch mode 
from a 1 m$^{3}$ IBC tank filled to 0.5 m$^{3}$ from the secondary treatment clarifier tank
at the FZ-J WWTP. Water was exchanged by completely draining the IBC tank into a
prior treatment tank then refilling with 0.5 m$^{3}$ of influent. The inner walls and
bottom of the IBC tank were scrubbed during water exchange to reduce algal growth
within the tank. Water exchange was performed Mondays through Thursdays, except
for Wednesday, October 3. Commercially available centrifugal sump pumps were used
to recirculate the water to the floway at a rate of 37 L min$^{-1}$ with a pulsing
interval of 15-20 s, provided by a tipping bucket at the headworks.  

*Harvesting and Biomass Processing*. Harvesting was performed weekly by turning off
the pumps, allowing the floway to drain for approximately 15 min, then manually 
scraping biomass from the base and walls of the lower 1.6 m of the floway, comprising
a total harvested area of 0.99 m$^2$. The remainder of the biomass was left to facilitate
re-colonization of the harvested area during the following growth interval. Scraped 
biomass was collected into a tared jar and weighed to determine the fresh weight 
of the harvested material. The biomass was dried by in an oven at 60°C for 24 hours
and weighed. Dried biomass was homogenized by grinding with a mortar and pestle. 
Weighed samples were frozen at -20° in airtight plastic bags.  

*Environmental Observations*. Floway water temperature and irradiance were logged 
once per minute using a HOBO data logger and downloaded as csv files using proprietary
software (Onset Computer Corporation, Bourne, MA).

*Water Chemistry*. One water sample (15 mL) was taken from the tailworks of the 
floway before and one after water exchange. These were either analyzed within six
hours or frozen for later analysis. Prior to analysis, samples were centrifuged 
at room temperature for 10 minutes at a relative centrifugal force of 3,007 g. Soluble
phosphate (PO$_4$-P) and Total nitrogen (TN) were determined using a DR 5000 UV-Vis
spectrophotometer after processing the samples with Hach LCK 349 (PO$_4$-P) and 
LCK 138 (TN) kits. (All: Hach Lange GmBH, Düsseldorf).  

*Biomass Composition*. Two 50 mg aliquots of each dried biomass sample, each with 2 
mL HNO$_3$ and 1 mL H$_2$O$_2$ were placed in closed Xpress containers (CEM Corporation, 
Matthews, NC, USA) and digested in a MARS 5 microwave digestion oven (CEM Corporation,
Matthews, NC, USA) by heating to 160°C within 20 minutes and kept at this temperature
for 15 minutes. The containers were then filled to 14 mL with deionized water. Two
aliquots of each sample solution obtained were diluted, one at 1:20 and one at 1:50,
and analyzed. Each dilution was assayed with ICP-OES (iCAP 6500, Thermo Scientific).
Reference material NIST1515 (SRM1515, 2017) was included as control. Recovery for
K was 101%, Ca 94%, Mg 94%, Mn 83% and P 90%. Measurements reflect the mean of two
determinations. Standard deviation for two observations, recorded originally as
"< 0,1" are set to NA in the cleaned data.  

# Data Cleaning  
To clean the water quality and biomass data, spreadsheets were loaded and reshaped
into a tidier schema before being saved as csv. Some columns of `ATS Treatment.xlsx`
("Dilution PO4-P", "PO4-P", "reduction", "Dilution TNb", "TNb", "remark"), were 
not retained in the cleaned data, while others (unnamed column with "before/after",
"commend") were converted from character to logical for improved utility. All columns
of `Analysen ZEA-3.xlsx` were retained. Temperature and irradiance logs (`ATS_<#.dd.mm.yy>.hobo`)
were converted to csv using proprietary software (`*.hproj` contain metadata about 
the `*.hobo` and were not processed). Because the data loggers were either initialized
in an office before being transported to the floway for installation, or downloaded 
after being removed from the floway and taken to an office, the initial and final 
observations of each file were discarded. Figures 1-7 and the code that generated 
them show which observations were discarded. Selected observations were combined 
into a single data frame and saved as a csv file.  

# HOBO Data.  
## Week 1 (ATS_1)  
```{r week1}
# Create cleaned data set from the first week.
ti_df <- CleanHOBOData(HOBO_fn[1], 1010, 365, w=1)
```

## Week 2 (ATS_2)  
```{r week2}
# Clean and append the data set from the following week.
ti_df <- ti_df %>% rbind(CleanHOBOData(HOBO_fn[2], 9773, 271, w=2))
```

## Week 3 (ATS_1)  
```{r week3}
# Clean and append the data set from the following week.
ti_df <- ti_df %>% rbind(CleanHOBOData(HOBO_fn[3], 119, 96, w=3))
```

## Week 4 (ATS_2)  
```{r week4}
# Clean and append the data set from the following week.
ti_df <- ti_df %>% rbind(CleanHOBOData(HOBO_fn[4], 107, 227, w=4))
```

## Week 5 (ATS_1)  
```{r week5}
# Clean and append the data set from the following week.
ti_df <- ti_df %>% rbind(CleanHOBOData(HOBO_fn[5], 107, 227, w=5))
```

## Week 6 (ATS_2)  
```{r week6}
# Clean and append the data set from the following week.
ti_df <- ti_df %>% rbind(CleanHOBOData(HOBO_fn[6], 121, 65, w=6))
```

## Week 7 (ATS_1)  
```{r week7}
# Clean and append the data set from the following week.
ti_df <- ti_df %>% rbind(CleanHOBOData(HOBO_fn[7], 84, 54, w=7))
```
```{r save_ti}
# Save the data set as a csv file
write_csv(ti_df, ti_fn)
```

# Floway Chemistry & Biomass Productivity
On two occasions, samples were assayed both on the day they were taken and after
freezing (Table 1). As the absolute differences between the readings for frozen
and unfrozen samples was small, less than 5% except for two measurements, it was
decided to freeze and accumulate samples for collective analysis rather than assay
each sample on the day it was collected.  

```{r WaterChemistryBiomass}
# Load and clean the data in "ATS Treatment.xlsx"
wcb_df <- WaterChemistryBiomass()

# Display data for samples assayed both frozen and fresh
kable(WaterChemistryBiomassDupes(), caption="Differences among fresh and frozen samples")

# Split the data into water chemistry and biomass productivity.
wc_df <- wcb_df %>%
    select(datetime, before, PO4P, TN, pH, frozen, assaydate, obsdate)

bp_df <- wcb_df %>%
    filter(!is.na(wet_biomass)) %>%
    select(wet_biomass, dry_biomass, obsdate)

# Save the water chemistry data, cleaned by removing duplicates.
wc_df <- wc_df[c(1:2, 4, 6:12, 14, 16:55),]
write_csv(wc_df, wc_fn)

# Save the cleaned biomass productivity data
write_csv(bp_df, bp_fn)
```

# Biomass Composition Data
```{r BiomassCompositionData}
# Load and clean the data in "Analysen ZEA-3.xlsx"
bc_df <- BiomassCompositionData(bcd)
write_csv(bc_df, bc_fn)

# Plot biomass productivity data
barplot(matrix(c(bp_df$dry_biomass, bp_df$wet_biomass), ncol=7, byrow=TRUE),
        xlab=bp_df$obsdate)

# Plot biomass composition data
PlotBiomassCompositionData(bc_df)
PlotMeanBiomassCompositionData(bc_df)
#PlotMolarRatios(bc_df)
```

# Data Dictionaries 
```{r bcdd}
# id
# date
# atom
# proportion
# sd
kable(data.frame(Column = colnames(bc_df),
                 Description = c("Observation ID",
                                 "Harvest date",
                                 "Element assayed",
                                 "Mass proportion of element in sample",
                                 "Standard deviation of proportion with n = 2")),
      caption = bc_fn)

# wet_biomass
# dry_biomass
# obsdate
kable(data.frame(Column = colnames(bp_df),
                 Description = c("Mass of total fresh biomass harvested",
                                 "Mass of total dried biomass",
                                 "Harvest date")),
      caption = bp_fn)

# datetime
# temp
# lux
# week
kable(data.frame(Column = colnames(ti_df),
                 Description = c("Date and time of observation",
                                 "Temperature (°C)",
                                 "Irradiance (lux)",
                                 "Week index")),
      caption = ti_fn)

# datetime
# before
# PO4P
# TN
# pH
# frozen
# assaydate
# obsdate
kable(data.frame(Column = colnames(wc_df),
                 Description = c("Date and time of observation",
                                 "Was this sample taken before or after water exchange?",
                                 "Soluble PO4, mg L-1",
                                 "Total N, mg L-1",
                                 "ph",
                                 "Was this sample frozen?",
                                 "Date sample was assayed",
                                 "Observation date")),
      caption = wc_fn)
```

# Acknowledgement
This work was supported by the German Federal Ministry of Education and Research
(BMBF) within the project Algnutrient-UrBioSol, Bioeconomy International FKZ031B0453A.

# References