<!-- Copyright 2019 by Forschungszentrum-Jülich (FZ-J). -->
<!-- This file is part of the publication "FZ-J Pilot ATS Floways" -->
<!-- (FZJPAF), documenting a concept for establishing a major ATS research -->
<!-- program at Forschungszentrum Jülich. The software component of FZJPAF is -->
<!-- open access: you can redistribute it and/or modify it under the terms of the -->
<!-- GNU General Public License as published by the Free Software Foundation, -->
<!-- either version 3 of the License, or (at your option) any later version. The -->
<!-- software component of FZJPAF is distributed in the hope that it will be -->
<!-- useful, but WITHOUT ANY WARRANTY; without even the implied warranty of -->
<!-- MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General -->
<!-- Public License for more details. You should have received a copy of the GNU -->
<!-- General Public License along with the software component of FZJPAF. If not, -->
<!-- see <http://www.gnu.org/licenses/>. -->

\tableofcontents
\newpage
\pagenumbering{arabic}  

```{r i_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "../../")
```
```{r i_settings, include=FALSE}
source("Code/R/Settings.R")
```
# Introduction
This document describes in detail the work needed to complete reconfiguration of the small ATS pilot floway at the FZ-J campus WWTP, and outlines additional work needed to complete the floways at the stabilization ponds and to implement floways in the Algae Park greenhouses. It does not address the large scale floways currently known as the Production Floways.

## Nomenclature
The original floway built in 2018 will henceforth be referred to as the "Exploratory Floways" (see below for explanation of the plural). The floways located adjacent to the stabilization basins at the WWTP will be referred to as the "WWTP Floways". The floways planned for the Algae Park location will be referred to as the "Novagreen Floways" and the "IVG Floways". These are all described in more detail below.

## General Requirements
Completing the planned floways will require several steps. First, floway components need to be constructed. This could be done off-site by a contractor and then shipped, or done on-site. Second, the components need to be assembled into completed floways. Students could participate in some of the construction and assembly part of their education. Figure 2 is a preview of drawings for the required components (still missing: dump bucket drawings). Third, any associated hardware such as plumbing, sensors, and gas sparging equipment needs to be fitted and tested. Water then needs to begin flowing.
